---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r, echo = FALSE}
library(dplyr)
library(ggplot2)
```

## Loading and preprocessing the data
```{r, echo = TRUE}
unzip("activity.zip")
df <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

1. Analysis of total steps per day
```{r, echo = TRUE}
daily_steps <- df %>% 
group_by(date) %>% 
summarise(steps_per_day = sum(steps, na.rm = TRUE))
```

2. Histogram of daily plots
```{r, echo = TRUE}
ggplot(daily_steps,
       aes(steps_per_day)) +
  geom_histogram(binwidth=500) + 
  xlab("Number of steps per day") +
  ylab("Frequency") +
  ggtitle("Original dataset")
```


3. The mean and median of the total number of steps taken per day
```{r, echo = TRUE}
summary(daily_steps)
```

## What is the average daily activity pattern?

1. Analysis of mean steps per interval
```{r, echo = TRUE}
interval_steps <- df %>% 
group_by(interval) %>% 
summarise(steps_per_interval = mean(steps, na.rm = TRUE))
```

2. Time series plot
```{r, echo = TRUE}
ggplot(data = interval_steps,
       aes(x = interval,
           y = steps_per_interval)) +
  geom_line() +
  xlab("Time intervals") +
  ylab("Average number of steps")
```


3. Interval with maximum number of steps
```{r, echo = TRUE}
interval_steps %>%
  arrange(desc(steps_per_interval)) %>% 
  head(n=1)
```
## Imputing missing values
1. Analysis of missing data
```{r, echo = TRUE}
lapply(df, function(x) sum(is.na(x)))
```
2. Imputing missing values by its mean
```{r, echo = TRUE}
df_imputed <- df
df_imputed$steps[is.na(df_imputed$steps)] <- mean(df_imputed$steps, na.rm = TRUE)
```


3. Analysis of total steps per day in imputed data
```{r, echo = TRUE}
daily_steps_imputed <- df_imputed %>% 
group_by(date) %>% 
summarise(steps_per_day = sum(steps))
```

4. Histogram of daily plots based on imputed dataset
```{r, echo = TRUE}
ggplot(daily_steps_imputed,
       aes(steps_per_day)) +
  geom_histogram(binwidth=500) + 
  xlab("Number of steps per day") +
  ylab("Frequency") +
  ggtitle("Imputed dataset")
```


5. Comparison of mean and median number of steps in original and imputed dataset

Original dataset:
```{r, echo = TRUE}
library(broom)
summary_original_df <- daily_steps$steps_per_day %>%
  summary() %>%
  tidy()
summary_original_df
```
Imputed dataset:
```{r, echo = TRUE}
summary_imputed_df <- daily_steps_imputed$steps_per_day %>%
  summary() %>%
  tidy()
print(summary_imputed_df)
```

```{r, echo = TRUE}
summary_imputed_df$mean - summary_original_df$mean
```
Mean of imputed data is higher by ~1412 steps than mean of original data

```{r, echo = TRUE}
summary_imputed_df$median - summary_original_df$median
```
Median of imputed data is higher by ~371 steps than median of original data

In conclusion imputation eliminated missing steps data and in consequence lead to increase of number of steps.


## Are there differences in activity patterns between weekdays and weekends?

1. Preparing new variable
```{r, echo = TRUE}
library(lubridate)

df_imputed$date <- as.Date(df_imputed$date)
df_imputed$weekday_type <- ifelse(wday(df_imputed$date, week_start = 1) >= 6, "weekend", "weekday")

df_imputed <- df_imputed %>% 
  group_by(interval, weekday_type) %>% 
  summarise(mean_steps = mean(steps))
```

2. Preparing the plot
```{r, echo = TRUE}
ggplot(df_imputed, aes(interval, mean_steps)) + 
  geom_line() + 
  facet_wrap(~weekday_type, nrow = 2) +
  xlab("5-Minute intervals") + 
  ylab("Average number of steps")
```
Weekdays have a more intense mornings, while steps over the weekend are more balanced